---
title: "R Notebook"
output: html_notebook
---

```{r}
library(tidyverse)
library(corrplot)
library(caTools)
library(Metrics)
```


```{r}
citiesAttributes <- read.csv("city_attributes.csv")
citiesHumidity <- read.csv("humidity.csv") 
citiesPressure <- read.csv("pressure.csv")
citiesTemp <- read.csv("temperature.csv")
citiesWeatherDescription <- read.csv("weather_description.csv")
citiesWind_direction <- read.csv("wind_direction.csv")
citiesWind_speed <- read.csv("wind_speed.csv")
```

```{r}
citiesHumiditySF <- select(citiesHumidity,"datetime" ,"San.Francisco")
names(citiesHumiditySF)[2] <- "Humidity"
citiesPressureSF <- select(citiesPressure ,"San.Francisco")
names(citiesPressureSF)[1] <- "Pressure"
citiesTempSF <- select(citiesTemp ,"San.Francisco")
names(citiesTempSF)[1] <- "Temperature"
citiesWeatherDescriptionSF <- select(citiesWeatherDescription ,"San.Francisco")
names(citiesWeatherDescriptionSF)[1] <- "Weather Description"
citiesWind_directionSF <- select(citiesWind_direction ,"San.Francisco")
names(citiesWind_directionSF)[1] <- "Wind Direction"
citiesWind_speedSF <- select(citiesWind_speed ,"San.Francisco")
names(citiesWind_speedSF)[1] <- "WindSpeed"
sanFranciscoWeatherData <- cbind(citiesHumiditySF,citiesPressureSF, citiesTempSF, citiesWeatherDescriptionSF, citiesWind_directionSF,citiesWind_speedSF)

sanFranciscoWeatherData <- sanFranciscoWeatherData[-c(1), ]
sanFranciscoWeatherData <- na.omit(sanFranciscoWeatherData)

```

```{r}
# Creating correlation Matrix and displaying it
correlation <- cor(sanFranciscoWeatherData[-c(1,5)])
corrplot(correlation, method="number")
```

```{r}
# Set the random seed for repeatability
set.seed(123) 

# Split the data into 3:1 ratio
sample = sample.split(sanFranciscoWeatherData, SplitRatio = .75)
trainingSet = subset(sanFranciscoWeatherData, sample == TRUE)
testingSet  = subset(sanFranciscoWeatherData, sample == FALSE)
```


# Build Model #1 on Training Set
```{r}
model <- lm(Temperature ~ Pressure, trainingSet)
model
```

# Predicting the testing set Model 1
```{r}
# Predict the position with the model and test data
predictedTemperatureWPressure <- predict(model, testingSet)

# Create data frame with actual and predicted values
temperaturePrediction <- data.frame(Pressure = testingSet$Pressure, 
                               actualTemperature = testingSet$Temperature,
                               predictedTemperature = predictedTemperatureWPressure)

temperaturePrediction
```

## Visualize the prediction Model 1
```{r}
ggplot(data = temperaturePrediction, mapping = aes(x = Pressure)) +
  geom_point(data = temperaturePrediction, 
             mapping = aes(y = actualTemperature),
             color = "blue") +
  geom_point(data = temperaturePrediction, 
             mapping = aes(y = predictedTemperature), 
             color = "red") +
  labs(title = "Pressure vs. Temperature" ,
        subtitle = "Actual (Blue) vs. Predicted (Red)", x = "Pressure",
        y = "Temperature")

```

# Mean Absolute Error Model 1

```{r}
error <- mse(temperaturePrediction$actualTemperature, temperaturePrediction$predictedTemperature)
cat("Mean Square Error:", error)
```

# Model 2

# Build Model #2 on Training Set
```{r}
model2 <- lm(Temperature ~ Humidity, trainingSet)
model2
```

# Predicting the testing set Model 2
```{r}
# Predict the position with the model and test data
predictedTemperatureWHumidity <- predict(model2, testingSet)

# Create data frame with actual and predicted values
HumidityPrediction <- data.frame(Humidity = testingSet$Humidity, 
                               actualTemperature = testingSet$Temperature,
                               predictedTemperature = predictedTemperatureWHumidity)


HumidityPrediction
```

## Visualize the prediction Model 2
```{r}
ggplot(data = HumidityPrediction, mapping = aes(x = Humidity)) +
  geom_point(data = HumidityPrediction, 
             mapping = aes(y = actualTemperature),
             color = "blue") +
  geom_point(data = HumidityPrediction, 
             mapping = aes(y = predictedTemperature), 
             color = "red") +
  labs(title = "Humidity vs. Temperature" ,
        subtitle = "Actual (Blue) vs. Predicted (Red)", x = "Pressure",
        y = "Temperature")

```

# Mean Absolute Error Model 2

```{r}
error <- mse(HumidityPrediction$actualTemperature, HumidityPrediction$predictedTemperature)
cat("Mean Square Error:", error)
```

# Model 3

# Build Model #3 on Training Set
```{r}
model3 <- lm(Temperature ~ WindSpeed, data = trainingSet)
model3
```

# Predicting the testing set Model 3
```{r}
# Predict the position with the model and test data
predictedTemperatureW_Wind_Speed <- predict(model3, testingSet)

# Create data frame with actual and predicted values
WindSpeedPrediction <- data.frame(WindSpeed = testingSet$WindSpeed, 
                               actualTemperature = testingSet$Temperature,
                               predictedTemperature = predictedTemperatureW_Wind_Speed)

WindSpeedPrediction
```

## Visualize the prediction Model 3
```{r}
ggplot(data = WindSpeedPrediction, mapping = aes(x = WindSpeed)) +
  geom_point(data = WindSpeedPrediction, 
             mapping = aes(y = actualTemperature),
             color = "blue") +
  geom_point(data = WindSpeedPrediction, 
             mapping = aes(y = predictedTemperature), 
             color = "red") +
  labs(title = "Wind Speed vs. Temperature" ,
        subtitle = "Actual (Blue) vs. Predicted (Red)", x = "Wind Speed",
        y = "Temperature")

```

# Mean Absolute Error Model 3

```{r}
error <- mse(WindSpeedPrediction$actualTemperature, WindSpeedPrediction$predictedTemperature)
cat("Mean Square Error:", error)
```

